### 2014-05-07 ApTranscriptome_TR.Rmd
After evaluating the effect of removing contaminants prior to quantifying gene expresssion, determined best course of action was to quantify expression for all genes and then remove the contaminants. This strategy ensures that reads associated with contaminants are removed during the quantification step.
Updated script to follow this quantification strategy. In process, updated to use Sailfish v0.6.3
which required some updates to the "sailfish quant" step.

**April 29, 2014**

@@

# Load Bioconductor libraries
#source("http://bioconductor.org/biocLite.R")
#biocLite("topGO")
#biocLite("Rgraphviz")
library(topGO) 
#suppressMessages(library(Rgraphviz))

@@

~~~
# download filtered Trinity assembly, uncompress and move
wget http://johnstantongeddes.org/assets/files/Aphaenogaster_transcriptome.tar
tar -xvf Aphaenogaster_transcriptome.tar
# check md5sum
md5sum Trinity_cap3_uclust_clean.fa
# c604159ee944dddda7b4e2a556bf2f53
mkdir -p results/
mkdir -p results/trinity-full/
mv Trinity_cap3_uclust_clean.fasta results/trinity-full/.
~~~

@@

I quantified gene expression using [sailfish](http://www.cs.cmu.edu/~ckingsf/software/sailfish/index.html). To run this program, first make sure that PATHs to the software libraries are set up correctly:

    export LD_LIBRARY_PATH=/opt/software/Sailfish-0.6.2-Linux_x86-64/lib:$LD_LIBRARY_PATH
    export PATH=/opt/software/Sailfish-0.6.2-Linux_x86-64/bin:$PATH

@@

    sailfish index -t Trinity_cap3_uclust_clean.fasta -o sailfish-index-Trinity_cap3_uclust -k 20 -p 4

@@

    mkdir sailfish-expression-Trinity-cap3-uclust

Then, for each sample, run the following command:

    sailfish -i sailfish-index -o sailfish-expression/A22-0 --reads A22-0_ATCACG.paired.left.fastq A22-0_ATCACG.paired.right.fastq A22-0_ATCACG.unpaired.left.fastq A22-0_ATCACG.unpaired.right.fastq -p 4

@@

```{r sailfish, eval=TRUE, echo=TRUE, cache=TRUE}
# directory containing trimmed reads
readdir <- "data/ind_files/" 
# sailfish index directory
sfindex <- "results/trinity-full/sailfish-index-Trinity_cap3_uclust"
# sailfish expression directory
sfexpressionroot <- "results/trinity-full/sailfish-expression-Trinity-cap3-uclust/"

@@

for (j in 1:length(samples)) {
    message("Start expression quantification for sample ", samples[j], ": ", Sys.time())
    quantdir <- paste(sfexpressionroot, samples[j], "_quant", sep="")
    samp.pos <- grep(paste(paste(samples[j], "_", sep="")), paired.left)
    samp.paired.l <- paste(readdir, paired.left[samp.pos], sep="")
    samp.paired.r <- paste(readdir, paired.right[samp.pos], sep="")
    samp.unpaired.l <- paste(readdir, unpaired.left[samp.pos], sep="")
    samp.unpaired.r <- paste(readdir, unpaired.right[samp.pos], sep="")
    sailfishcmd <- paste("sailfish quant -i ", sfindex, " -o ", quantdir, " --reads ", samp.paired.l, " ", samp.paired.r, " ", samp.unpaired.l, " ", samp.unpaired.r, " -p 4", sep="")
#   print(sailfishcmd)
    system(sailfishcmd)
    message("Done with expression quantification for sample ", samples[j], ": ", Sys.time(), "\n")
}
```

@@

```{r load_expression_data, eval=TRUE, echo=FALSE, results='hide'}
# read in each file using loop
samples <- c("A22-0", "A22-3", "A22-7", "A22-10", "A22-14", "A22-17", "A22-21", "A22-24", "A22-28", "A22-31", "A22-35", "A22-38", "Ar-0", "Ar-3", "Ar-7", "Ar-10", "Ar-14", "Ar-17", "Ar-21", "Ar-24", "Ar-28", "Ar-31", "Ar-35", "Ar-38")

for (j in 1:length(samples)) {
    samp <- samples[j]
    trtval <- as.numeric(str_split_fixed(samp, "-", 2)[2])
    outpre <- gsub("-", "_", samp)
    outname <- paste(outpre, "_quant", sep="")
    read.sailfish.quant(filein=paste(sfexpressionroot, samp, "_quant/quant_bias_corrected.sf", sep=""), outname=outname, samp = samp, trtval = trtval)
}

@@

```{r exp_correlations, echo=FALSE, results='asis'}
# Make Rmarkdown pandoc table of correlations among expression levels between colonies at each temp

temp <- c(0, 3.5, 7, 10.5, 14, 17.5, 21, 24.5, 28, 31.5, 35, 38.5)
cors <- c(round(cor(Ar_0_quant$TPM, A22_0_quant$TPM), 2), round(cor(Ar_3_quant$TPM, A22_3_quant$TPM), 2), round(cor(Ar_7_quant$TPM, A22_7_quant$TPM), 2), round(cor(Ar_10_quant$TPM, A22_10_quant$TPM), 2), round(cor(Ar_14_quant$TPM, A22_14_quant$TPM), 2), round(cor(Ar_17_quant$TPM, A22_17_quant$TPM), 2), round(cor(Ar_21_quant$TPM, A22_21_quant$TPM), 2), round(cor(Ar_24_quant$TPM, A22_24_quant$TPM), 2), round(cor(Ar_28_quant$TPM, A22_28_quant$TPM), 2), round(cor(Ar_31_quant$TPM, A22_31_quant$TPM), 2), round(cor(Ar_35_quant$TPM, A22_35_quant$TPM), 2), round(cor(Ar_38_quant$TPM, A22_38_quant$TPM), 2))

cortable <- cbind(temp, cors)

pandoc.table(cortable, style="rmarkdown", caption = "correlations between colonies at each temperature treatment")
```

## Identification of thermally-responsive genes

>>>>

**May 6, 2014**

@@

# Load Bioconductor libraries
#source("http://bioconductor.org/biocLite.R")
#biocLite("topGO")
#biocLite("Rgraphviz")
library(topGO) 
library(Rgraphviz)

@@

~~~
# download filtered Trinity assembly, uncompress and move
wget http://johnstantongeddes.org/assets/files/Aphaenogaster_transcriptome.tar
tar -xvf Aphaenogaster_transcriptome.tar
# check md5sum
md5sum Trinity_cap3_uclust_clean.fa
# 0160a82ec62c6f371c84e4963fed09d1
mkdir -p results/
mkdir -p results/trinity-full/
mv Trinity_cap3_uclust.fasta results/trinity-full/.
mv Trinity_cap3_uclust_clean.fasta results/trinity-full/.
~~~

@@

I quantified gene expression using [sailfish](http://www.cs.cmu.edu/~ckingsf/software/sailfish/index.html). To run this program, first make sure that PATHs to the software libraries are set up correctly as described on the sailfish website.

@@

```{r sailfish_index}
system("sailfish index -t results/trinity-full/Trinity_cap3_uclust.fasta -o results/trinity-full/sailfish-index-Trinity-cap3-uclust -k 20 -p 4")
```

@@

```{r}
system("mkdir -p results/trinity-full/sailfish-expression-Trinity-cap3-uclust")
```

Then, for each sample, run the following command in the `results/trinity-full/` directory:

    sailfish -i sailfish-index-Trinity-cap3-uclust -o sailfish-expression-Trinity-cap3-uclust/A22-0 -l "T=PE:O=><:S=U" -r A22-0_ATCACG.unpaired.left.fastq A22-0_ATCACG.unpaired.right.fastq -1 A22-0_ATCACG.paired.left.fastq -2 A22-0_ATCACG.paired.right.fastq  -p 4

@@

```{r sailfish, eval=TRUE, echo=TRUE, cache=TRUE}
# directory containing trimmed reads
readdir <- "data/ind_files/" 
# sailfish index directory
sfindex <- "results/trinity-full/sailfish-index-Trinity-cap3-uclust"
# sailfish expression directory
sfexpressionroot <- "results/trinity-full/sailfish-expression-Trinity-cap3-uclust/"

@@

for (j in 1:length(samples)) {
    message("Start expression quantification for sample ", samples[j], ": ", Sys.time())
    quantdir <- paste(sfexpressionroot, samples[j], "_quant", sep="")
    samp.pos <- grep(paste(paste(samples[j], "_", sep="")), paired.left)
    samp.paired.l <- paste(readdir, paired.left[samp.pos], sep="")
    samp.paired.r <- paste(readdir, paired.right[samp.pos], sep="")
    samp.unpaired.l <- paste(readdir, unpaired.left[samp.pos], sep="")
    samp.unpaired.r <- paste(readdir, unpaired.right[samp.pos], sep="")
    sailfishcmd <- paste("sailfish quant -i ", sfindex, " -o ", quantdir, " -l 'T=SE:S=U' -r ", samp.paired.l, " ", samp.paired.r, " ", samp.unpaired.l, " ", samp.unpaired.r, " -p 4", sep="")
    #print(sailfishcmd)
    system(sailfishcmd)
    message("Done with expression quantification for sample ", samples[j], ": ", Sys.time(), "\n")
}
```

@@

```{r load_expression_data, eval=TRUE, echo=FALSE, results='hide'}
# read in each file using loop
for (j in 1:length(samples)) {
    samp <- samples[j]
    trtval <- as.numeric(str_split_fixed(samp, "-", 2)[2])
    outpre <- gsub("-", "_", samp)
    outname <- paste(outpre, "_quant", sep="")
    read.sailfish.quant(filein=paste(sfexpressionroot, samp, "_quant/quant_bias_corrected.sf", sep=""), outname=outname, samp = samp, trtval = trtval)
}

@@

```{r exp_correlations, echo=FALSE, results='hide'}
temp <- c(0, 3.5, 7, 10.5, 14, 17.5, 21, 24.5, 28, 31.5, 35, 38.5)
cors <- c(round(cor(Ar_0_quant$TPM, A22_0_quant$TPM), 2), round(cor(Ar_3_quant$TPM, A22_3_quant$TPM), 2), round(cor(Ar_7_quant$TPM, A22_7_quant$TPM), 2), round(cor(Ar_10_quant$TPM, A22_10_quant$TPM), 2), round(cor(Ar_14_quant$TPM, A22_14_quant$TPM), 2), round(cor(Ar_17_quant$TPM, A22_17_quant$TPM), 2), round(cor(Ar_21_quant$TPM, A22_21_quant$TPM), 2), round(cor(Ar_24_quant$TPM, A22_24_quant$TPM), 2), round(cor(Ar_28_quant$TPM, A22_28_quant$TPM), 2), round(cor(Ar_31_quant$TPM, A22_31_quant$TPM), 2), round(cor(Ar_35_quant$TPM, A22_35_quant$TPM), 2), round(cor(Ar_38_quant$TPM, A22_38_quant$TPM), 2))

cortable <- cbind(temp, cors)
```

```{r exp_correlations_table, echo=FALSE, results='asis'}
pandoc.table(cortable, style="rmarkdown", caption = "correlations between colonies at each temperature treatment")
```

### Regression-model to identify thermally-responsive genes
